---
title: "Getting Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting-Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ggplot2)
library(ggmulticomp)


df = iris |>
  tidyr::pivot_longer(-Species )

base_plot = df |>
  ggplot(aes(x = Species, y = value)) +
  geom_boxplot() +
  facet_wrap(vars(name), scales = "free")

base_plot

```


# Tukey Test


Simply add `stat_tukey()` layer to your ggplot object. It will run Tukey test and show statistical significance for you. You will have to play around with `ggplot2::scale_y_continuous()` so the significance marker is present.

```{r}
base_plot +
  stat_tukey() +
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.5)))

```


You can use the `p.format` argument to change how p-values are presented. Two common options are `scales::label_pvalue()` or `scales::label_number_auto()`. 

```{r}
base_plot +
  stat_tukey(p.format = scales::label_pvalue()) +
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.5)))

```

You can also implement your own function to format it. This function should accept double as input, and return a character to be displayed.

```{r}

my_format = function(x) {
  dplyr::case_when(
    x < 0.001 ~ "very significant",
    x < 0.05 ~ "significant",
    TRUE ~  "not significant"
  )
}


base_plot +
  stat_tukey(p.format = my_format) +
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.5)))

```


By default, `stat_tukey` perform statistical tests on the values that are being presented in ggplot. That is, if you add `scale_y_log2()` to your ggplot, the statistic results will be performed on the log2-transformed y-values. If this is not a desired behavior, the `trans_value` argument allow you to apply a transformation function to the y-values before running the statistical test.

```{r}

base_plot +
  stat_tukey(p.format = scales::label_scientific()) + 
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.5))) +
  labs(title = "No y transformation")

base_plot +
  stat_tukey(trans_value = log2, p.format = scales::label_scientific()) + 
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.5)))  +
  labs(title = "Y values transformed before stat test")

```


You can fine tune some of the aesthetic parameters in case they do not look good by default.
```{r}
base_plot +
  stat_tukey(text_size = 5, line_size = 0.5, linetype = 2) + 
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.5)))

```

# Dunnett

Dunnett test works similar to Tukey test. The parameter should work similarly. One key distinction is that `stat_dunnett()` shows the p-values at the top of the group

```{r}
base_plot +
  stat_dunnett()
```

A different style of alignment can be used.

```{r}
base_plot +
  stat_dunnett(p.y_align = 2)

```

If you want explicitly show which comparisons are made, use `GeomBracket` as the geom.

```{r}

base_plot +
  stat_dunnett(geom = "bracket") + 
  scale_y_continuous(expand = expansion(mult = c(0.1, 0.5)))

```
